---
title: "Predict Citibike Rides"
output: html_notebook
---

First load in the data from the 

```{r}
library(tidyverse)
library(lubridate)

trips_per_day <- read.table('trips_per_day.tsv', header = TRUE)
trips_per_day


#Updating my dataframe to include additional features 
#Feature for month, day and and weekend 
library(lubridate)
trips_per_day %>% select(ymd, date)

# this creates columns to indicate day of week, month and if it's a weekend or not 
trips_per_day <-  
  trips_per_day %>%
  mutate(month = as.factor(month(ymd)), day_of_week = as.factor(wday(ymd))) %>%
  mutate(weekend = ifelse(day_of_week == 1 | day_of_week == 7, 1, 0)) %>% 
  mutate(is_rain = ifelse(prcp > 0.5, 1, 0)) %>%
  mutate(temp_diff = tmax - tmin) %>%
  mutate(is_snwd = ifelse(snwd > 5, 1, 0)) %>%
  mutate(avg_temp = (tmin+tmax)/2) 

trips_per_day

```

After doing data wrangling outside, joining the holidays file 
```{r}
library(readr)
holidays <- read_delim("holidays.csv", "\t", 
    escape_double = FALSE, col_names = FALSE) 

holidays <- 
  holidays %>%
  rename(
    holiday_num = X1, 
    ymd = X2, 
    holiday = X3 
  )

holidays <- 
  holidays %>%
  mutate(ymd = as.character(ymd))

holidays

```
Join the holiday data and create a column for whether a day is a holiday or not 
```{r}
trips_per_day <- 
  trips_per_day %>% 
  left_join(holidays)

trips_per_day

# create a column for whether a day is a holiday or not 
trips_per_day <-
  trips_per_day %>%
  mutate(holiday = ifelse(!is.na(holiday), 1, 0))

trips_per_day

```



-------------------------------------------------
## Split the data 

Now split the data 
80% for training the model 
10% for validation 
10% for final test set 
Reference used: https://stackoverflow.com/questions/36068963/r-how-to-split-a-data-frame-into-training-validation-and-test-sets

```{r}
set.seed(2)

num_days <- nrow(trips_per_day)
# set out the the fractions that you need to train
frac_train <- 0.8
frac_validation <- 0.1
frac_test <- 0.1 

# compute sample sizes 
sample_size_train <- floor(num_days * frac_train)
sample_size_validation <- floor(num_days * frac_validation)
sample_size_test <- floor(num_days * frac_test)

# randomly sample rows for the training set without replacement - he named this ndx
indices_traning <- sample(1:num_days, sample_size_train, replace = FALSE)

# setdiff() used to avoid overlapping subsets of indices 
indices_not_training <- setdiff(seq_len(num_days), indices_traning)

# now to find the validation and test indices 
indices_validation <- sample(indices_not_training, size = sample_size_validation)
indices_test <- setdiff(indices_not_training, indices_validation)


# used to fit the model 
trips_per_day_train <- trips_per_day[indices_traning, ]
trips_per_day_validate <- trips_per_day[indices_validation, ]

# Restricted access - don't look!
trips_per_day_test <- trips_per_day[indices_test, ]

# we can see the rest of the data
trips_per_day_train
trips_per_day_validate



```
_________________________________________________________
## Let's get to modeling

Start out with the model in that notebook, which uses only the minimum temperature on each day to predict the number of trips taken that day. Try different polynomial degrees in the minimum temperature and check that you get results similar to what's in that notebook, although they likely won't be identical due to shuffling of which days end up in the train, validation, and test splits. Quantify your performance using root mean-squared error.

Now we try plottign minimum temperature with the number of trips for different polynomial degrees

```{r}
# fitting a model for each polynomial degree
degree <- 1:8

# set the vectors to hold the training and validation errors
train_err <- c()
validate_err <- c()

for(deg in degree){
  
  # fit on training data 
  model <- lm(num_trips ~ poly(tmin, deg, raw =TRUE), data = trips_per_day_train)
  
  #par(mfrow= c(2,2))
  #plot(model)
  
  # using mean squared error
  # evaluate on training data  
  train_err[deg] <- sqrt(mean((predict(model, trips_per_day_train) - trips_per_day_train$num_trips)^2))
  
  #evaluate on validate data
  validate_err[deg] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))


}

train_err
validate_err


```

Now to plot the two errors together 
Note for me:
Gather takes multiple columns and collapses into key-value pairs, duplicating all other columns as needed. You use gather() when you notice that you have columns that are not variables.
```{r}
# this takes the degrees and separates the training and validation error 
plot_data <- 
  data.frame(degree, train_err, validate_err) %>% 
  gather("split", "error", -degree)

# Now let's plot 
plot_data %>%
  ggplot(aes(x=degree, y= error, color = split)) +
  geom_line() +
  scale_x_continuous(breaks = degree) +
  xlab("Polynomial degree") +
  ylab("RMSE")

  
```

Now to fit the data and the model together. I choose the 4th degree polynomial to stay on the safe side and not do overfitting

```{r}
library(dplyr)
library(modelr)
library(broom)

model <- lm(num_trips ~ poly(tmin, 4, raw = T), data = trips_per_day_train)

trips_per_day_train <- 
  trips_per_day_train %>%
  add_predictions(model) %>%
  mutate(split = "train")

trips_per_day_validate <- 
  trips_per_day_validate %>%
  add_predictions(model) %>% 
  mutate(split = "validate")

plot_data <- bind_rows(trips_per_day_train, trips_per_day_validate)


plot_data %>%
ggplot(aes(x = tmin, y = num_trips)) +
  geom_point(aes(color = split)) +
  geom_line(aes(y = pred)) +
  xlab('Minimum temperature') +
  ylab('Daily trips') +
  scale_y_continuous()


trips_per_day_train
rmse(model, plot_data)


```

```{r}
tidy(model)
glance(model)
```


Let's see what we're working with I keep forgetting

```{r}
colnames(trips_per_day)
```


----------------------------------------------------

## Weather model: Model 1: This model looks at effects of weather alone with precipitation, snow, maximum temperature and minimum temperature
```{r}

weather_model <- lm(num_trips ~ prcp*snwd + tmax*tmin, trips_per_day_train)
summary(weather_model)
rmse(weather_model, trips_per_day_validate)

```
To evaluate the performance of the model. 
```{r}
trips_per_day_train <- 
  trips_per_day_train %>%
  add_predictions(weather_model) %>%
  mutate(split = "train")

trips_per_day_validate <- 
  trips_per_day_validate %>%
  add_predictions(weather_model) %>% 
  mutate(split = "validate")

plot_data <- bind_rows(trips_per_day_train, trips_per_day_validate)

plot_data %>%
  ggplot(aes(x=pred, y=num_trips)) +
  geom_point(aes(color=split)) +
  geom_abline(linetype = "dashed") +
  xlab("Predicted") +
  ylab("Actual")

# calculate the rmse - the values don't look too different
rmse(weather_model, trips_per_day_train)
rmse(weather_model, trips_per_day_validate)
rmse(weather_model, plot_data)

```

---------------------------------------------------------

## Days of the week model: Model 2:
How do the days of the week affect number of trips?
```{r}

trips_per_day_train %>% 
  ggplot(aes(x=day_of_week, y=num_trips)) +
  geom_point(aes(color = tmin))

```
Now let's create the model between weekend and taking tweaks from the previous model 
```{r}
#model_2 <- lm(num_trips ~ weekend + prcp*snwd + tmax*tmin, trips_per_day_train)
#summary(model_2)
# RMSE: 4545.876
# R-squared: 0.8604

model_02 <- lm(num_trips ~ weekend + prcp + snwd + tmax +tmin, trips_per_day_train)
summary(model_02)
# RMSE: 3895.621
# R-squared: 0.859

# modifying model_2 based on the the summary
trips_per_day_train <- 
  trips_per_day_train %>%
  add_predictions(model_02) %>%
  mutate(split = "train")

trips_per_day_validate <- 
  trips_per_day_validate %>%
  add_predictions(model_02) %>% 
  mutate(split = "validate")

plot_data <- bind_rows(trips_per_day_train, trips_per_day_validate)

plot_data %>%
  ggplot(aes(x=pred, y=num_trips)) +
  geom_point(aes(color=split)) +
  geom_abline(linetype = "dashed") +
  xlab("Predicted") +
  ylab("Actual")

# calculate the rmse - the values don't look too different
rmse(model_02, plot_data)
rmse(model_02, trips_per_day_validate) #4122.867

```

-----------------------------------------------------------
Model 03: Now let's use the simpler model and testing with the different polynomials for precipitation

```{r}
# fitting a model for each polynomial degree
degree <- 1:8

# set the vectors to hold the training and validation errors
train_err <- c()
validate_err <- c()

for(deg in degree){
  
  # fit on training data 
  model_03 <- lm(num_trips ~ poly(prcp, deg, raw =TRUE) + weekend + tmax + snwd + tmin, data = trips_per_day_train)

  
  # using mean squared error
  # evaluate on training data  
  train_err[deg] <- sqrt(mean((predict(model_03, trips_per_day_train) - trips_per_day_train$num_trips)^2))
  
  #evaluate on validate data
  validate_err[deg] <- sqrt(mean((predict(model_03, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))

}

# this takes the degrees and separates the training and validation error 
plot_data <- 
  data.frame(degree, train_err, validate_err) %>% 
  gather("split", "error", -degree)

# Now let's plot 
plot_data %>%
  ggplot(aes(x=degree, y= error, color = split)) +
  geom_line() +
  scale_x_continuous(breaks = degree) +
  xlab("Polynomial degree for Precipitation") +
  ylab("RMSE")



```


Let's stay with polynomial degree 3 for precipitation to stay on the safe side for overfitting 
```{r}
model_03 <- lm(num_trips ~ poly(prcp, 3, raw =TRUE) + weekend + tmax + snwd + tmin, data = trips_per_day_train)

# modifying model_02 based on the the summary
trips_per_day_train <- 
  trips_per_day_train %>%
  add_predictions(model_03) %>%
  mutate(split = "train")

trips_per_day_validate <- 
  trips_per_day_validate %>%
  add_predictions(model_03) %>% 
  mutate(split = "validate")

plot_data <- bind_rows(trips_per_day_train, trips_per_day_validate)

plot_data %>%
  ggplot(aes(x=pred, y=num_trips)) +
  geom_point(aes(color=split)) +
  geom_abline(linetype = "dashed") +
  xlab("Predicted") +
  ylab("Actual")

summary(model_03)
rmse(model_03, plot_data)
rmse(model_03, trips_per_day_validate)

# RMSE: 3607.688
# R-squared: 0.8778
```
With this slightly more complicated model, we're getting a slightly better RMSE 
Model_03:
RMSE: 3607.688
R-squared: 0.8878
But the problem is, it takes the polynomial of prcp so it might be prone to overfitting 

--------------------------------------------------------

Model 04: Tweak of the model before but accounting day of the week and whether it was a holiday or not 
```{r}
model_04 <- lm(num_trips ~ weekend + tmax +tmin + day_of_week + holiday, trips_per_day_train)


# there's NA as a variable for day 6 which indicate dthat the variable in question is linearly related to the other variables. If this is the case, then there's no unique solution to the regression without dropping one of the variables

# now testing
trips_per_day_train <- 
  trips_per_day_train %>%
  add_predictions(model_04) %>%
  mutate(split = "train")

trips_per_day_validate <- 
  trips_per_day_validate %>%
  add_predictions(model_04) %>% 
  mutate(split = "validate")

plot_data <- bind_rows(trips_per_day_train, trips_per_day_validate)

plot_data %>%
  ggplot(aes(x=pred, y=num_trips)) +
  geom_point(aes(color=split)) +
  geom_abline(linetype = "dashed") +
  xlab("Predicted") +
  ylab("Actual")

summary(model_04)
rmse(model_04, plot_data)
rmse(model_04, trips_per_day_validate) #5318.257

# RMSE: 4774.246
# R-squared: 0.7914


```
Now how can we decide an amount of rain as substantial
```{r}

trips_per_day %>% 
  ggplot(aes(x=prcp, y=num_trips)) +
  geom_point(aes(color=tmin)) +
  facet_wrap(~ weekend) +
  geom_smooth(method = "lm")

```
Let's decide substantial precipitation is precipitation greater than 
```{r}
model_05 <- lm(num_trips ~  is_rain*weekend + tmax + tmin + day_of_week + holiday + month, trips_per_day_train)

summary(model_05)

# now testing
trips_per_day_train <- 
  trips_per_day_train %>%
  add_predictions(model_05) %>%
  mutate(split = "train")

trips_per_day_validate <- 
  trips_per_day_validate %>%
  add_predictions(model_05) %>% 
  mutate(split = "validate")

plot_data <- bind_rows(trips_per_day_train, trips_per_day_validate)

plot_data %>%
  ggplot(aes(x=pred, y=num_trips)) +
  geom_point(aes(color=split)) +
  geom_abline(linetype = "dashed") +
  xlab("Predicted") +
  ylab("Actual")

summary(model_05)
rmse(model_05, plot_data)
rmse(model_05, trips_per_day_validate) #4462.155

# RMSE: 3555.428
# R_squared: 0.8892

```


--------------------------------------------------------
Model 06:
Does the ratio or difference of tmax and tmin have an effect?
```{r}
trips_per_day_train %>% 
  mutate(temp_ratio = tmax/tmin) %>%
  ggplot(aes(x=temp_ratio, y=num_trips)) +
  geom_point() 

trips_per_day_train %>% 
  mutate(temp_ratio = tmax-tmin) %>%
  ggplot(aes(x=temp_ratio, y=num_trips)) +
  geom_point() +
  scale_x_log10() +
  geom_smooth(method="lm") 

```

There does not seem to be a direct correlation so I'm not going to incorporate. But let's try it out with the temperature difference
```{r}
model_06 <- lm(num_trips ~ is_rain*weekend + tmin + temp_diff + month + holiday + day_of_week, trips_per_day_train)

# now testing
trips_per_day_train <- 
  trips_per_day_train %>%
  add_predictions(model_06) %>%
  mutate(split = "train")

trips_per_day_validate <- 
  trips_per_day_validate %>%
  add_predictions(model_06) %>% 
  mutate(split = "validate")

plot_data <- bind_rows(trips_per_day_train, trips_per_day_validate)

plot_data %>%
  ggplot(aes(x=pred, y=num_trips)) +
  geom_point(aes(color=split)) +
  geom_abline(linetype = "dashed") +
  xlab("Predicted") +
  ylab("Actual")

summary(model_06)
rmse(model_06, plot_data)
rmse(model_06, trips_per_day_validate) #4462.155

```

Does daily trips differ ny month?
```{r}
trips_per_day_train %>% 
  ggplot(aes(x=month, y=num_trips)) +
  geom_point() +
  geom_smooth(method = "lm")
```
Yep, using month was a reasonable choice

-------------------------------------------------------
Model 7: This would be model 2 but with days of the week and the month
```{r}
model_07 <- lm(num_trips ~ weekend*is_rain + is_snwd  + day_of_week + month*weekend + holiday + tmin + tmax, trips_per_day_train)

summary(model_07)
rmse(model_07, trips_per_day_validate)
```

--------------------------------------------------------
Model 8: Using the polynomial data 

```{r}
model_08 <- lm(num_trips ~ poly(prcp, 3, raw=T) + month + day_of_week + tmin + tmax + is_snwd*weekend + holiday, trips_per_day_train)

# modifying model_02 based on the the summary
trips_per_day_train <- 
  trips_per_day_train %>%
  add_predictions(model_08) %>%
  mutate(split = "train")

trips_per_day_validate <- 
  trips_per_day_validate %>%
  add_predictions(model_08) %>% 
  mutate(split = "validate")

plot_data <- bind_rows(trips_per_day_train, trips_per_day_validate)

plot_data %>%
  ggplot(aes(x=pred, y=num_trips)) +
  geom_point(aes(color=split)) +
  geom_abline(linetype = "dashed") +
  xlab("Predicted") +
  ylab("Actual")

summary(model_08)
rmse(model_08, plot_data)
rmse(model_08, trips_per_day_validate)

```

-----------------------------------------------------------
Model 9: Model that uses the average temperature 

```{r}
trips_per_day_train %>%
  ggplot(aes(x=avg_temp, num_trips)) +
  geom_point()+
  geom_smooth(method = "lm")
```


```{r}
# fitting a model for each polynomial degree
degree <- 1:8

# set the vectors to hold the training and validation errors
train_err <- c()
validate_err <- c()

for(deg in degree){
  
  # fit on training data 
  model <- lm(num_trips ~ poly(avg_temp, deg, raw =TRUE), data = trips_per_day_train)
  
  #par(mfrow= c(2,2))
  #plot(model)
  
  # using mean squared error
  # evaluate on training data  
  train_err[deg] <- sqrt(mean((predict(model, trips_per_day_train) - trips_per_day_train$num_trips)^2))
  
  #evaluate on validate data
  validate_err[deg] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))


}

train_err
validate_err


# this takes the degrees and separates the training and validation error 
plot_data <- 
  data.frame(degree, train_err, validate_err) %>% 
  gather("split", "error", -degree)

# Now let's plot 
plot_data %>%
  ggplot(aes(x=degree, y= error, color = split)) +
  geom_line() +
  scale_x_continuous(breaks = degree) +
  xlab("Polynomial degree") +
  ylab("RMSE")

```

So let's create a final model
```{r}
model_09 <- lm(num_trips ~ avg_temp + poly(prcp, 3, raw=T) + weekend + is_snwd + holiday , trips_per_day_train)


# modifying model_02 based on the the summary
trips_per_day_train <- 
  trips_per_day_train %>%
  add_predictions(model_09) %>%
  mutate(split = "train")

trips_per_day_validate <- 
  trips_per_day_validate %>%
  add_predictions(model_09) %>% 
  mutate(split = "validate")

plot_data <- bind_rows(trips_per_day_train, trips_per_day_validate)

plot_data %>%
  ggplot(aes(x=pred, y=num_trips)) +
  geom_point(aes(color=split)) +
  geom_abline(linetype = "dashed") +
  xlab("Predicted") +
  ylab("Actual")

summary(model_09)
#rmse(model_09, plot_data)
rmse(model_09, trips_per_day_train)
rmse(model_09, trips_per_day_validate)

```



_______________________________________________________

## Choosing a model  
```{r}
model_list <- list(weather_model, model_02, model_03, model_04, model_05, model_06, model_07, model_08, model_09)

train_err <- c()
validate_err <- c()

for(model in model_list){
  
  train_err <- c(train_err, sqrt(mean((predict(model, trips_per_day_train) - trips_per_day_train$num_trips)^2)))
  
  #evaluate on validate data
  validate_err <- c(validate_err, sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2)))


}

train_err
validate_err

models <- seq(1:length(model_list))
plot_data <- 
  data.frame(models, train_err, validate_err) %>% 
  gather("split", "error", -models)

plot_data %>%
  ggplot(aes(x=models, y= error, color = split)) +
  geom_line() +
  scale_x_continuous(breaks = models) +
  xlab("Model") +
  ylab("RMSE")


```

________________________________________________________________________
## Plot the final model in two different ways
I am choosing Model 9 to be my final model 
Now to plot the model in two different ways 
```{r}
# modifying model_02 based on the the summary
trips_per_day_train <- 
  trips_per_day_train %>%
  add_predictions(model_09) %>%
  mutate(split = "train")

trips_per_day_validate <- 
  trips_per_day_validate %>%
  add_predictions(model_09) %>% 
  mutate(split = "validate")

plot_data <- bind_rows(trips_per_day_train, trips_per_day_validate)
plot_data <- 
  plot_data %>%
  add_predictions(model_09)

# now convert integer date to date format for the plot
plot_data <- transform(plot_data, date = as.Date(as.character(date), "%Y%m%d"))

plot_data %>%
  ggplot(aes(x=date)) +
  geom_point(aes(y=num_trips, color = split)) +
  geom_smooth(aes(y=pred), size = 2, se=F) +
  scale_x_date() +
  xlab("Date") +
  ylab("Number of Trips")


```
Now a plot with the predicted value on the x-axis and the actual value on the y-axis
```{r}
plot_data %>%
  ggplot(aes(x=pred, y=num_trips)) +
  geom_point(aes(color=split)) +
  geom_abline(linetype = "dashed") +
  xlab("Predicted") +
  ylab("Actual")

```

_____________________________________________________________________
## Save the final model!
```{r}

# create a function for your data transformation

save(model_09, file = 'final_model.RData')


```

___________________________________________
## Now testing on test data 


```{r}

rmse(model_09, trips_per_day_test)


trips_per_day_test <- 
  trips_per_day_validate %>%
  add_predictions(model_09)  


trips_per_day_test %>%
  ggplot(aes(x=pred, y=num_trips)) +
  geom_point(aes(y=num_trips)) +
  geom_abline(linetype = "dashed") +
  xlab("Predicted") +
  ylab("Actual")

```


Results:
The RMSE on the test data is 3232.411
And the actual and predicted line is a diagonal linear line which is ideal.









