---
title: "Testing CitiBike Predictions"
output: html_notebook
---

## Let's load the trip data
```{r}
load('trips_2015.RData')

# Get rid of ymd and add date
weather <- 
  weather %>%
  select(-ymd)

weather <- 
  weather %>% 
  rename(
    ymd = date
  ) 


```

Now let's see what trips and weather look like
```{r}
trips
colnames(trips)
colnames(weather)

trips <- 
  trips %>%
  transform(
    ymd = as.character(ymd)
  )

weather 
trips 

trips %>%
  select(ymd)

```
Let's calculate num_trips 
```{r}

trips <- 
  trips %>%
  select(ymd) %>%
  group_by(ymd) %>%
  summarise(num_trips = n())

trips


```


Let's create columns to contain what I had for the original working data 

```{r}
weather

trips_per_day <- 
  trips %>% 
  left_join(weather)

trips_per_day %>%
  transform(
    tmax = tmax/10, 
    tmin = tmin/10
  )



```

Now joining the holidays data
Let's import holidays and join that 
```{r}

holidays <- read_delim("holidays.csv", "\t", 
    escape_double = FALSE, col_names = FALSE) 

holidays <- 
  holidays %>%
  rename(
    holiday_num = X1, 
    ymd = X2, 
    holiday = X3 
  )


holidays <- 
  holidays %>%
  mutate(ymd = as.character(ymd))


trips_per_day <- 
  trips_per_day %>% 
  left_join(holidays, by = "ymd")

# create a column for whether a day is a holiday or not 
trips_per_day <-
  trips_per_day %>%
  mutate(holiday = ifelse(!is.na(holiday), 1, 0))


trips_per_day

```





```{r}

# now to modify it to include my features
trips_per_day <-  
  trips_per_day %>%
  mutate(month = as.factor(month(ymd)), day_of_week = as.factor(wday(ymd))) %>%
  mutate(weekend = ifelse(day_of_week == 1 | day_of_week == 7, 1, 0)) %>% 
  mutate(is_rain = ifelse(prcp > 0.5, 1, 0)) %>%
  mutate(is_snwd = ifelse(snwd > 5, 1, 0)) %>%
  mutate(avg_temp = (tmin+tmax)/2)

```


Let's see if the data has the features I want 
```{r}
trips_per_day

```
_________________________________________________

## Now to test my model on the data 

Compute the RMSE between the actual and predicted trips for 2015 and compare the results to what you found with cross-validation on the 2014 data.

```{r}
model <- lm(num_trips ~ avg_temp + poly(prcp, 3, raw=T) + weekend + is_snwd + holiday, trips_per_day)

rmse(model, trips_per_day)


trips_per_day <- 
  trips_per_day %>%
  add_predictions(model_09)  

trips_per_day %>%
  ggplot(aes(x=pred, y=num_trips)) +
  geom_point(aes(y=num_trips)) +
  geom_abline(linetype = "dashed") +
  xlab("Predicted") +
  ylab("Actual")




```
The predictions were very bad. 
The RMSE is 6806.518 and the predicted line is very far away. This has made me sad. 
The predictions for 2014 had RMSE on the test data to be 3232.411. Which was similar train data and validate data 
And the actual and predicted line was diagonal linear line through the points which was ideal.

_______________________________________________

## Now waiting to record test results for a teammmate




_______________________________________________
## Final Thoughts














