---
title: "LAB 3.6.2 ISLR"
output: html_notebook
---
Import the libraries

```{r}
library(MASS)
library(ISLR)
```

```{r}
#fix(Boston)
names(Boston)
?Boston

```

```{r}
attach(Boston)
lm.fit=lm(medv~lstat)
summary(lm.fit)
names(lm.fit)
```
To extract coefficients 
```{r}
coef(lm.fit)
```
```{r}
# the confidence interval 
confint(lm.fit)

# produce confidence intervals 
predict (lm.fit ,data.frame(lstat=c(5,10 ,15)),interval ="confidence")

# produce prediction intervals
predict(lm.fit, data.frame(lstat=c(5, 10, 15)), interval="prediction")



```
Now let's get plotting!
```{r}
# lwd is for line thickness
plot(lstat, medv)
abline(lm.fit)
abline (lm.fit ,lwd =3)
abline (lm.fit ,lwd=3 ,col ="red")


```

```{r}
plot(lstat ,medv,lwd = 5 ,col="blue")
plot(lstat ,medv ,pch =20)
plot(1:20,1:20,pch =1:20)


```

This plot is too cool for school
```{r}
plot(lstat ,medv ,pch ="😎")

```

par() is used to divide the plotting region
```{r}
par(mfrow=c(2,2))
plot(lm.fit)

```

Residuals from linear regression fit and rstident()
```{r}
plot(predict(lm.fit), residuals(lm.fit))
plot(predict(lm.fit), rstudent(lm.fit))
```


Leverage statistics can be computed for any number of predictors using the hatvalues() function

```{r}
plot(hatvalues(lm.fit))
# now to tell us the largest leverage statistic 
which.max(hatvalues(lm.fit))
```



##############################################################


## Lab 3.6.3 throught 3.6.6

Lab 3.6.3 
```{r}
lm.fit = lm(medv~lstat+age, data=Boston)
summary(lm.fit)

summary(lm.fit)$sigma

```
```{r}
# install.packages("car")
# library(car)
# vif(lm.fit)
```

If want to do regression with everything except one variable 
```{r}
lm.fit
```

















