---
title: "ISRS Question Answers and Testing"
output: html_notebook
---
######################################

Exercises 2.2 and 2.5

######################################



Problem 2.2:
Of the 34 patients in the control group, 4 were alive at the end of the study. Of the 69 patients in
the treatment group, 24 were alive. The contingency table below summarizes these results.
(a) What proportion of patients in the treatment group and what proportion of patients in the
control group died?
```{r}

```


(b) One approach for investigating whether or not the treatment is effective is to use a random-
ization technique.

i. What are the claims being tested? Use the same null and alternative hypothesis notation
used in the section.

ii. The paragraph below describes the set up for such approach, if we were to do it with-
out using statistical software. Fill in the blanks with a number or phrase, whichever is appropriate.
We write alive on cards representing patients who were alive at the end of the study, and dead on cards representing patients who were not. Then, we shuffle these cards and split them into two groups:
one group of size representing treatment, and another group of size representing control. We calculate the difference between the proportion of dead cards in the treatment and control groups (treatment -
control) and record this value. We repeat this many times to build a distribution
centered at . Lastly, we calculate the fraction of simulations where
the simulated differences in proportions are . If this fraction is low,
we conclude that it is unlikely to have observed such an outcome by chance and
that the null hypothesis should be rejected in favor of the alternative.

iii. What do the simulation results shown below suggest about the effectiveness of the trans-
plant program?


########################################


Exercise 5.2 
For figure 1: the data does not have a direct linear trend. It has non-constant variability so it is not fit for linear regression. 

For figure 2: the data shows a linear trend with constant variability but is has however the residuals are not normal. Residuals are normal if plotting the residuals forms a straight line 

Book ans:
(a) The residual plot will show randomly
distributed residuals around 0. The variance is
also approximately constant. 
(b) The residuals
will show a fan shape, with higher variability for
smaller x. There will also be many points on the
right above the line. There is trouble with the
model being fit here.

Exercise 5.9:
1,302 UCLA students were asked to fill out a survey where they were
asked about their height, fastest speed they have ever driven, and gender. The scatterplot on the
left displays the relationship between height and fastest speed, and the scatterplot on the right
displays the breakdown by gender in this relationship.
(a) Describe the relationship between height and fastest speed. 
The relationship does not seem linear
(b) Why do you think these variables are positively associated?
(c) What role does gender play in the relationship between height and fastest driving speed?
Ans: 
(a) The relationship is positive, weak, and
possibly linear. However, there do appear to
be some anomalous observations along the left
where several students have the same height
that is notably far from the cloud of the other
points. Additionally, there are many students
who appear not to have driven a car, and they are represented by a set of points along the bot-
tom of the scatterplot. 
(b) There is no obvious explanation why simply being tall should lead a person to drive faster. However, one confounding factor is gender. Males tend to be taller than females on average, and personal experi-
ences (anecdotal) may suggest they drive faster. If we were to follow-up on this suspicion, we
would find that sociological studies confirm this
suspicion. 
(c) Males are taller on average and
they drive faster. The gender variable is indeed
an important confounding variable.

#################################################


Exercise 5.20: 
With the information:
a. 
```{r}
# All in cm 
mu_shoulder <- 108.20
sig_shoulder <- 10.37
mu_height <- 171.14
sig_height <- 9.41
R_shoulder_height <- 0.67

b <- ( R_shoulder_height * sig_shoulder )/ sig_height
b

a <- mu_height - b * mu_shoulder
a

# the equation: y = a + bx

```
The equation is y = 91.25023 + 0.7383528x

b. slope, b = 0.7383528
  intercept,a = 91.25023

c. Calculate R squared 
```{r}
R_squared <- R * R 
R_squared
```
The higher the value of R squared the better the model fits the data. A value of 0.5184 means that it might fit the data well about 50%. Half of the observed variation can be explained by the model's input. 

d. shoulder girth of 100cm 
```{r}
stu_height <-  (91.25023 + 0.7383528 *100)
stu_height
```
e. Residual = 165.0855 - 160 cm
```{r}
residual <- 165.0855 - 160
residual
```
Residual is the leftover variation in data after accounting for model fit 

f. Child has girth 56cm
```{r}
child_height <- (91.25023 + 0.7383528 * 56)
child_height

```

This is an anomaly so not modelled well with the data.


Exercise 5.29: 
Let's try to recreate the plot 
```{r}
library(tidyverse)
library(dplyr)
body_data <- read.table("body.dat.txt", header = FALSE)
# columns 23:weight and 24:height contain the weight and 
#weight_height_df <- body_data[,23:24]

#rename the columns as weight and height
weight_height_df <- weight_height_df %>% 
  rename(
    weight = V23, 
    height = V24 
)
weight_height_df

```
Let's try to recreate the plot from the question 
```{r}
library(ggplot2)
ggplot(weight_height_df, aes(x = height, y = weight))+
  geom_point(color = "#7E95F3") +
  xlab("Height (in cm)") +
  ylab("Weight (in kg)") +
  scale_y_continuous() +
  geom_smooth(method = "lm")

```
Reproduce the table: 
Some notes: 
The lm() function creates a regression model given some formula in the form of Y ~ X + X2
```{r}
library(scales)
library(broom)
library(modelr)


model <- lm(weight ~ height, weight_height_df)

summary(model)

```
Now answer the questions: 
a. Height and weight have a positive correlation. For a unit increase in weight, height increases by 1.0176
Things that I missed: 
The relationship is positive, moderate-to-strong and linear. There are a few outliers but no points that appear to be influential. 

b.Write the equation of the regression line. Interpret the slope and intercept in context.

Ans: weight = -105.0113 + height*1.0176  

c. Do the data provide strong evidence that an increase in height is associated with an increase
in weight? State the null and alternative hypotheses, report the p-value, and state your
conclusion.
Ans: The data provides strong evidence that an increase in height is associated with an increase in weight as the P-value that something like this would occur outside of this scenario is 0.0000. 
We can make it a one-sided test:
The null hypothesis, H0: The true slope of the coefficient of height is zero (beta = 0)
Alternative hypothesis, hA: The true slope of the coefficient of height is greater is zero (beta > 0)

#### The two sided hypothesis test:
The p-value for
the two-sided alternative hypothesis (beta 6= 0) is incredibly small, so the p-value for the one-sided hypothesis will be even smaller. That is, we reject H0. The data provide convincing evidence that height and weight are positively correlated. The true slope parameter is indeed
greater than 0.

c. R = 0.72
```{r}
R <- 0.72
R_squared <- 0.72 * 0.72
print(R_squared)
```

## Chapter 6 

#### Exercise 6.1, 6.2, 6.3

```{r}
baby_data <- read.table("babyweights.txt")
baby_data 
```

Now to produce the table 
```{r}

summary(weight_smoke_model)


```
Can we plot it to see what is looks like 
```{r}
ggplot(baby_data, aes(x = smoke, y = weight, color= smoke))+
  geom_point() +
  geom_smooth(method = "lm")
```

```{r}
weight_smoke_model <- lm(weight~smoke, baby_data)
summary(weight_smoke_model)
```


a. weight_hat = 129.6359 - 2.8122 * smoke 

b. For smokers
weight_hat = 129.6359 - 2.8122*1 = 126.8237
For non-smokers:
weight_hat = 129.6359 - 2.8122*0 = 129.6359

c. H0: beta = 0 
HA:beta =/= 0 
As the p value is 0.0236
The t-value is the estimate divided by their standard errors and the given t vaue is -2.267 and the p-value is 0.0236 so we reject the null hypothesis in favor of the alternative hypothesis. 


### Exercise 6.2 
```{r}
#Recreate the results in the example 
weight_parity_model <- lm(weight~parity, baby_data)
summary(weight_parity_model)
```
a. The equation of the regression line:
weight = 129.7390 - 4.2953 * parity 

b. For first borns :
weight = 129.7390 - 4.2953 * 0 = 129.7390
For otherwise:
weight = 129.7390 - 4.2953 * 1 = 125.4473

c. Ho: beta = 0 
Ha : beta =/= 0 
As the p-value is 0.00185 so we reject the null hypothesis in favor of the alternative hypothesis. 

### Exercise 6.3
Comparing all the values with weight 
```{r}
weight_model <- lm(weight~ . ,baby_data)
summary(weight_model)
par(mfrow = c(2,2))
plot(weight_model)
```
a. Equation of regression that might include all the variables 
weight_hat = -107.42291 + 0.06724*bwt + -0.02805*gestation -3.25319 * parity +  0.42734 * age + 3.52377 * height + -1.99493 * smoke 

b. The slope fo gestation is -0.02805 so as gestation increases by one unit, the weight decreases by -0.02805
As age increases by one unit, the weight increases by 0.42734

c. The coefficient for parity is different than in the linear model shown in Exercise 6.2. Why
might there be a difference?
Ans: Parity may be correlated with one of the other in the model, which complicates model estimation 

d. According to the first observation: 
```{r}
weight_hat <-   -107.42291 + 0.06724*120 + -0.02805* 284 -3.25319 * 0+  0.42734 * 27 + 3.52377 * 62  + -1.99493 * 0
weight_hat
weight_residual <- 100 -weight_hat
weight_residual
```
The model overpredicts the baby's weight by 22.69161 units


e. Calculation R_squared and adjusted R_squared 
```{r}
var_res <- 249.28
var_outcome <- 332.57
R_squared <- 1 - (var_res/var_outcome)
R_squared

# calculate the adjusted value 
num_of_cases <- 1236
num_of_predictors <-  ncol(baby_data) - 1
R_squared_adj <- R_squared * (num_of_cases -1)/(num_of_cases - num_of_predictors -1 )
R_squared_adj
```

























