---
title: "IST Notebook Question Answers Chp 4,6,7,9,10"
output: html_notebook
---

Question 4.1 
Value Probability
0         p
1         2p
2         3p
3         4p
4         5p
5         6p
Table 4.4: The Distribution of Y

Table 4.6 presents the probabilities of the random variable Y
These probabilities are a function of the number p, the probability of the value
“0”. Answer the following questions:
1. What is the value of p?
Ans: p = 1/21 
2. P(Y < 3) = 6p = 6/21
3. P(Y = odd) = 12p = 12/21
4. P(1 ≤ Y < 4) = 9p = 9/21 = 3/7 
5. P(|Y − 3| < 1.5) = 15p = 12/21 = 4/7
6. E(Y ) = 
The calculation involves multiplying the probabilities with the values
```{r}
Y.val <- c(0,1,2,3,4,5)
P.val <- c(1,2,3,4,5,6)/21
E <- sum(Y.val*P.val)
print(E)

```

7. Var(Y ) = ?
The calc involves 
```{r}
Var <- sum((Y.val-E)^2*P.val)
Var

```

8. What is the standard deviation of Y
```{r}
sd <- sqrt(Var)
sd
```

Some testing from pages in the chapter: 
```{r}
pnorm(5,2,3) - pnorm(0,2,3)
qnorm(0.75)

# checking the qnorm function
qnorm(0.025, 2, 3)

#checking out pbinom
pbinom(2060, 4000, 0.5) - pbinom(1939, 4000, 0.5)
```




Question 6.1 : Consider the problem of establishing regulations concerning the
maximum number of people who can occupy a lift. In particular, we would like
to assess the probability of exceeding maximal weight when 8 people are allowed
to use the lift simultaneously and compare that to the probability of allowing 9
people into the lift.
Assume that the total weight of 8 people chosen at random follows a normal
distribution with a mean of 560kg and a standard deviation of 57kg. Assume
that the total weight of 9 people chosen at random follows a normal distribution
with a mean of 630kg and a standard deviation of 61kg.
1. What is the probability that the total weight of 8 people exceeds 650kg?
```{r}
1 - pnorm(650, 560, 57)
```

2. What is the probability that the total weight of 9 people exceeds 650kg?
```{r}
1 - pnorm(650, 630, 61)
```

3. What is the central region that contains 80% of distribution of the total
weight of 8 people?
The 80% range would be between 90% and 10%
```{r}
qnorm(0.1, 560, 57)
qnorm(0.9, 560, 57)
```
4. What is the central region that contains 80% of distribution of the total
weight of 9 people?
```{r}
qnorm(0.1, 630, 61)
qnorm(0.9, 630, 61)


# question: couldn't you just do 80%?
# Because they specified it to be the central region
qnorm(0.8, 630, 61)
```
####################################
Exercise 7.1 
Our goal in this question is to investigate the sampling distribution of the sample
average of the variable “bmi”. We assume a sample of size n = 150.
1. Compute the population average of the variable “bmi”.
```{r}
directory <- ("~//Desktop/")
# this part isn't reading in the file 
#pop.2 <- read.csv(file = directory)

# trying another way 
pop.2 <- read.csv("http://pluto.huji.ac.il/~msby/StatThink/Datasets/pop2.
csv")

# CONTINUING WITH THE ASSUMPTION THAT THE FILE IS BEING PROPERLY BEING READ 
mean(pop.2$bmi)
```

2. Compute the population standard deviation of the variable “bmi”.
```{r}
sd(pop.2$bmi)
```

3. Compute the expectation of the sampling distribution for the sample average of the variable.
Answer: 
Initially, we produce a vector of zeros of the given lenght (100,000). In each
iteration of the “for” loop a random sample of size 150 is selected from the population. The sample average is computed and stored in the sequence “X.bar”.
At the end of all the iterations all the zeros are replaced by evaluations of the
sample average.
The expectation of the sampling distribution of the sample average is computed by the application of the function “mean” to the sequence that represents
the sampling distribution of the sample average.
```{r}
# this produces a vector of zeros size of 10^5 and populate it with zeros 
X.bar <- rep(0, 10^5)

for(i in 1:10^5)
{
  # take a 150 person sample from the population
  X.samp <- sample(pop.2$bmi, 150)
  # then calculate the mean of the sample and store in the vector 
  X.bar[i] <- mean(X.samp) 
}

# After that's done take the mean of the sample
mean(X.bar)
mean_val <- mean(X.bar)
```

4. Compute the standard deviation of the sampling distribution for the sample average of the variable.
```{r}
sd(X.bar)
sd_val <- sd(X.bar)
```

5. Identify, using simulations, the central region that contains 80% of the
sampling distribution of the sample average.
```{r}
# here we can just use quantile to find the central region that contains 80% of the distribution
quantile(X.bar, c(0.1, 0.9))

```

6. Identify, using the Central Limit Theorem, an approximation of the central region that contains 80% of the sampling distribution of the sample
average.

```{r}
qnorm(0.1, mean_val, sd_val)
qnorm(0.9, mean_val, sd_val)

# Another way to do this 
qnorm(c(0.1, 0.9), mean_val, sd_val)
```



###########################################
Exercise 9.1: A total of 50 patients experiencing post-polio pain syndrome were recruited.
Some of the patients were treated with an active magnetic device and the others
were treated with an inactive placebo device. All patients rated their pain before
(score1) and after application of the device (score2). The variable “change”
is the difference between “score1” and “score2. The treatment condition is
indicated by the variable “active.” The value “1” indicates subjects receiving
treatment with the active magnet and the value “2” indicates subjects treated
with the inactive placebo.

Answer: 
1.What is the sample average of the change in score between the patient’s
rating before the application of the device and the rating after the application?

```{r}
# Import the data 
magnets <- read.csv("http://pluto.huji.ac.il/~msby/
StatThink/Datasets/magnets.csv")

summary(magnets)

# now to find the mean of the change 
mean(abs(magnets$score2 - magnets$score1))

```
2. Is the variable “active” a factor or a numeric variable?
```{r}
typeof(magnets$active)
magnets$active
```
We see that it is a character type. 

3. . Compute the average value of the variable “change” for the patients that
received and active magnet and average value for those that received an
inactive placebo. (Hint: Notice that the first 29 patients received an active
magnet and the last 21 patients received an inactive placebo. The subsequence of the first 29 values of the given variables can be obtained via
the expression “change[1:29]” and the last 21 vales are obtained via the
expression “change[30:50]”.)
```{r}
treatment_mean <- mean(magnets$change[1:29], na.rm = TRUE)
non_treatment_mean <- mean(magnets$change[30:50])

treatment_mean
non_treatment_mean

```

4. Compute the sample standard deviation of the variable “change” for the
patients that received and active magnet and the sample standard deviation for those that received an inactive placebo.

```{r}
sd(magnets$change[1:29])
sd(magnets$change[30:50])
```

5. Produce a boxplot of the variable “change” for the patients that received
and active magnet and for patients that received an inactive placebo.
What is the number of outliers in each subsequence?
```{r}
library(ggplot2)
boxplot(magnets$change[1:29])
boxplot(magnets$change[30:50])
```
There are three outliers in the non-treatment boxplot. 
There are no outliers in the treatment boxplot. 
We can look at the outliers as: 
```{r}
table(magnets$change[30:50])
```
But when we the table of values we see that there are actually 4 outliers which are 3,4,5,11



######################################################
Checking out some code from Chapter 10 
```{r}
lam <- 1/13000
 X.bar <- rep(0,10^5)
 for(i in 1:10^5)
 {
 X <- rexp(201,lam)
 X.bar[i] <- mean(X)
 }
 mean(abs(X.bar - 1/lam) <= 1000)
```

Prodice a sampling distribution 
```{r}
mu <- 3
sig <- sqrt(2)
X.bar <- rep(0, 10^5)
mid.range <- rep(0,10^5)
for(i in 1:10^5)
{
  X <- rnorm(100, mu, sig)
  X.bar[i] <- mean(X)
  mid.range[i] <- (max(X) + min(X))/2
}
var(X.bar)
var(mid.range)
mean(X)
mean(X.bar)
mean(mid.range)


hist(X.bar)
hist(mid.range)
```
```{r}
 mu <- 5
 std <- sqrt(3)
 X.var <- rep(0,10^5)
 for(i in 1:10^5)
 {
 X <- rnorm(200,mu,std)
 X.var[i] <- var(X)
 }
 mean(X.var)
 
 hist(X.var)
```

######################################################

Exercise 10:
1. Simulate the sampling distribution of average and the median of a sample
of size n = 100 from the Normal(3, 2) distribution. Compute the expectation and the variance of the sample average and of the sample median.
Which of the two estimators has a smaller mean square error?
```{r}
summary(X)
hist(X)
view(X)

mu <- 3
sig <- sqrt(2)
X.bar <- rep(0, 10^5)
X.med<- rep(0, 10^5) 

for(i in 1:10^5)
{
  X <- rnorm(100, mu, sig)
  X.bar[i] <- mean(X)
  X.med[i] <- median(X)
}

# for sample mean
mean(X.bar)
var(X.bar)

# for sample median 
mean(X.med)
var(X.med)

#actual values 
mean(X)
```
The mean square error of an unbiased estimator is equal to its expectation

Note: my values are not exactly matching, what could be 
the error ?


2. Simulate the sampling distribution of average and the median of a sample
of size n = 100 from the Uniform(0.5, 5.5) distribution. Compute the
expectation and the variance of the sample average and of the sample
median. Which of the two estimators has a smaller mean square error?

```{r}
X.bar <- rep(0, 10^5)
X.med <- rep(0, 10^5)

for(i in 1:10^5)
{
  X <- runif(100, 0.5, 5.5)
  X.bar[i] <- mean(X)
  X.med[i] <- median(X)
}

mean(X.bar)
mean(X.med)

var(X.bar)
var(X.med) 

#what's the real mean 
mean(X)

# let's see what they look like 
hist(X.bar)
hist(X.med)
```
Exercise 10.2: 
 One of the variables in
the file is a factor by the name “group” that identifies levels of blood pressure.
The levels of this variable are “HIGH”, “LOW”, and “NORMAL”.
```{r}
# importing data again for sake of compeltion 
pop.2 <- read.csv("http://pluto.huji.ac.il/~msby/StatThink/Datasets/pop2.
csv")

ex2 <- read.csv("http://pluto.huji.ac.il/~msby/StatThink/Datasets/ex2.csv")

summary(ex2)
```

1. Compute the proportion in the sample of those with a high level of blood
pressure
```{r}
colnames(ex.2)
summary(ex.2$group)
high_group <- ex2 %>% select(group) %>% filter(group =="HIGH")%>% summarize(count = n()) 
high_group

# here we see the count is 37 
print(37/150)

summary(ex2)

# now to calculate the mean of the group of high
mean(ex2$group == "HIGH")

```
2. Compute the proportion in the population of those with a high level of
blood pressure.

```{r}
#pop2 <- read.csv("http://pluto.huji.ac.il/~msby/StatThink/Datasets/pop2.csv")

# calculate mean by group HIGH
mean(pop2$group == "HIGH")
```

3. Simulate the sampling distribution of the sample proportion and compute
its expectation.
```{r}

P.hat <- rep(0, 10^5)
for(i in 1:10^5)
{
  X <- sample(pop2$group, 200)
  P.hat[i] <- mean(X == "HIGH")
}
mean(P.hat)

```


4. Compute the variance of the sample proportion.
```{r}
var(P.hat)
```

5. It is proposed in Section 10.5 that the variance of the sample proportion
is Var(Pˆ) = p(1 − p)/n, where p is the probability of the event (having a
high blood pressure in our case) and n is the sample size (n = 150 in our
case). Examine this proposal in the current setting.

```{r}
# first calculate the probability of it happening 
p <- mean(pop.2$group == "HIGH")
p*(1-p)/150
```


# Exercise(s) is from Intro to Stat with Randomization and Simulation













 




